### 2016-10-14 mcmc.tex
got rid of a DFM

One note to make here
  is that you must \emph{propose}
  in the same coordinates (parameterization or transformation of parameters)
  as that in which your priors are \emph{specified},
  or else multiply in a (possibly nasty) Jacobian.
That is, if your prior is ``flat in $\pars$''
  but you find it easier to run your sampler by proposing steps in $\ln \pars$,
  if you don't modify your acceptance probability by the Jacobian,
  your \emph{real} prior won't be flat in $\pars$.
These think-os can get subtle;
  the best way to avoid them
  is to write your code and your likelihood function
  and your prior pdf function and your proposal distribution
  all in precisely the same parameterization.
We will return to this point below in our comments on testing (\sectionname~\ref{sec:trouble}).

@@

HOGG SAY HERE: ...You must describe your priors in the same parameterization that you ``step'' in.
That is, if you change your parameters, you \emph{must} change your priors correspondingly.
Give the example of going from $A$, $\phi$ to $A\,\cos\phi$, $A\,\sin\phi$.

@@

\paragraph{visualization:}
HOGG: What plots do you make to diagnose problems?

HOGG: Plot parameters vs ``time'' in the chain.

HOGG: Plot a cut through the density function and the samples near that cut.

HOGG:...Make triangle plots.  Great value!...  Show example here... Mentioned above; drop here or reiterate?

@@

DFM: SHOULD WE DO THIS ALL EXPLICITLY HERE IN MATH?
Choose a $p(A)$ and a $p(\phi) = [2\pi]^{-1}$,
   and then transform to $p(a, b)$.

%% ...Priors: Once again:
%% If you describe your priors in one parameterization but
%% then perform sampling in a different space, it is easy to make
%% mistakes.  For example: Flat prior in amplitude and phase, sample in
%% $A\,\cos\phi$ and $A\,\sin\phi$.  There are many ways to do this
%% wrong.

@@

HOGG: ...How to deal with discrete or integer or binary parameters?

>>>>

One note to make here
  is that you must \emph{propose}
  in the same coordinates (parameterization or transformation of parameters)
  as that in which your priors are \emph{specified},
  or else multiply in a (possibly nasty) Jacobian.
That is, if your prior is ``flat in $\pars$''
  but you find it easier to run your sampler by proposing steps in $\ln \pars$,
  if you don't modify your acceptance probability by the Jacobian,
  your \emph{real} prior won't be flat in $\pars$.
These think-os can get subtle;
  the best way to avoid them
  is to write your code and your likelihood function
  and your prior pdf function and your proposal distribution
  all in precisely the same parameterization.
We will return to this point below in our comments on testing (\sectionname~\ref{sec:trouble}).
The point is that if you \emph{do} change coordinates between
  the statement of your priors and the specific variables you
  step in, you have to introduce Jacobian factors.

@@

@@

@@

To be explicit, imagine that you have parameters $A$ and $\phi$ and
priors $p(A)$ and $p(\phi)$.
Now imagine that instead (and very sensibly) you want to sample in
parameters $x\equiv A\,\cos(\phi)$ and $y\equiv A\,\sin(\phi)$.
What is the proper prior to apply to $x, y$?
It is not just $p(A)\,p(\phi)$ with $A=\sqrt{x^2 + y^2}$ and $\phi=\arctan(y,x)$!
That's close, though! There is a correction factor which is the
absolute value of the determinant of the Jacobian matrix
$||\frac{\dd(A,\phi}{\dd(x,y)}||$.
That is, it is the determinant of the full derivative of one set of
parameters with respect to the other.
Keep your units straight!\note{See \citet{hoggcalculus} for advice on
  how to check and test these transformations using dimensional
  analysis.}
In this particular case, the Jacobian is equal to $A$, such that 
\begin{eqnarray}
  p(x,y) &=& \frac{1}{A}\,p(A)\,p(\phi)
  \quad ,
\end{eqnarray}
where $A=\sqrt{x^2 + y^2}$ and $\phi=\arctan(y,x)$.\note{Exercise to
  the reader: Do the same but with $x\equiv \sqrt{A}\,\cos(\phi)$ and
  $y\equiv\sqrt{A}\,\sin(\phi)$.}

@@

For another example, sometimes some parameters are continuous, and some
are integer (or discrete).
In this case, it is sufficient, in the context of M-H \MCMC, to make
custom proposal distributions for the integer parameters that make sure
only integers are proposed (provided care is taken to make sure that detailed
balance is preserved).
An example is to have a proposal that has some probability of
incrementing the parameter up, and an equal (ie, the same) probability
of incrementing it down.
Some samplers implicitly or explicitly assume that all parameters are
continuous\note{Once again, \citet{emcee}.}; these should be avoided
when some parameters are integer.
If you \emph{really} want to use such samplers when some parameters
are discrete, you have to put a \code{round()} or \code{int()}
operation into the likelihood and prior functions.
This makes for a stepwise-constant posterior pdf. It is a hack, but it
works in most circumstances.
